---
title: "Homework08"
author: "Mark D. Mullinger"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Creating fake data sets to explore hypotheses

Create a random data set with assigned sample size, mean, and standard deviation:
``` {r }
library(ggplot2)
library(tidyverse)
fxn <- function(n, mean, sd) {as.vector(mean + sd * scale(rnorm(n)))
}

v1 <- fxn(1000, 75, 17) 
v2 <- fxn(1000, 30, 17)
v3 <- fxn(1000, 5, 17) 
df <- data.frame(v1, v2, v3)
dflong <- df%>%
  pivot_longer(cols = 1:3,names_to = "Group",values_to = "Response")

head(dflong)
```

``` {r }

test <- aov(Response ~ Group, data=dflong)
summary(test)
```

``` {r}
p1 <- ggplot(data=dflong) +
               aes(x=Group,y=Response) + 
               geom_boxplot()
print(p1)
```

# Now change the means and find the threshhold for statistical significance

``` {r }
v1 <- fxn(1000, 75, 17) 
v2 <- fxn(1000, 75.5, 17)
v3 <- fxn(1000, 74.5, 17) 
df <- data.frame(v1, v2, v3)
dflong <- df%>%
  pivot_longer(cols = 1:3,names_to = "Group",values_to = "Response")

test <- aov(Response ~ Group, data=dflong)
summary(test)
```


```{r }
p1 <- ggplot(data=dflong) +
               aes(x=Group,y=Response) + 
               geom_boxplot()
print(p1)
```

The only way I could make p>0.05 was by making the means very close, only 0.5 difference from the 'original' mean of 75. Even means of 74, 75, and 76 showed significant differences with this sample size and variance. 
