---
title: "Homework10"
author: "Mark D. Mullinger"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Homework 10 is a practice of using for loops and randomization tests on data.

1. The first question required constructing a numeric vector with a number of zeros in it, and then using a for loop, write code that will count the number of zeros:
``` {r }
y <- c(0,1,2,4,5,6,7,0,0,5,4,3,5,5,67,4,3,0,2,0,3,4,5,6,0,3)
func1<-function(x){
  counter <- 0
for (i in x) {
  if(i == 0) counter <- counter + 1
}
print(counter)
  
}
func1(y)
```

2. Use subsetting to achieve the same result:
``` {r }
counter = sum(y==0)
print(counter)
```

3. Write a function that takes an input of two integers, and generates an output with each element as the result of row number * column number. I can change these inputs by calling the function with different variables.
``` {r }
mafxn <- function(nrow=4,ncol=12) {
  m <- matrix(nrow=nrow, ncol=ncol)
  for (i in 1:nrow){
    for (j in 1:ncol){
      m[i,j] = i*j
    }
  }
  return(m)
}
mafxn()
```
4. 

