---
title: "Homework07"
author: "Mark D. Mullinger"
date: "2023-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Run the sample code:
```{r }
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

```{r }
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```


2. Try it with your own data:
```{r }
z <- read.table("Phyllostachys.csv",header=TRUE,sep=",")
str(z)
summary(z)
```


```{r }
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

```{r }
# plot a histogram
p1 <- ggplot(data=z, aes(x=floret, y=after_stat(density))) +
  geom_histogram(color="grey60",fill="cornsilk",linewidth=0.2) 
print(p1)
```

```{r }
# add empirical density curve
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

```{r }
# get ML parameters for normal
normPars <- fitdistr(z$floret,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

```{r }
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$floret),len=length(z$floret))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$floret), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

```{r }
# plot normal probability density 
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$floret),len=length(z$floret))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$floret), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

```{r }
# plot exponential probability density
expoPars <- fitdistr(z$floret,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$floret), args = list(rate=rateML))
 p1 + stat + stat2
```

```{r }
# plot uniform probability density
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$floret), args = list(min=min(z$floret), max=max(z$floret)))
p1 + stat + stat2 + stat3
```

```{r }
# plot gamma probability density
gammaPars <- fitdistr(z$floret,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$floret), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

```{r }
# plot beta probability density. needs seperate plot because data must be rescaled
pSpecial <- ggplot(data=z, aes(x=floret/(max(floret + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")
pSpecial

betaPars <- fitdistr(x=z$floret/max(z$floret + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$floret), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

3. Find the best fitting distribution: I believe either the gamma or the beta curve fit these data the best. The lack of an upper bound as described in this question perhaps would make the gamma curve the best fitting.

4. Simulate a new data set with the same parameters as the original set: 
mean = 16.2641509. sd = 9.9939937

```{r}
z <- rnorm(n=318,mean=16.2641509,sd=9.9939937)
z <- data.frame(1:318,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

```{r}
# plot histogram of data
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

```{r}


gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]
xval <- seq(0,max(z$myVar),len=length(z$myVar))
stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat4
```

I think the simulated data did a decent job of simulating realistic measurements of floret number in this bamboo species. The values which make up the majority of measuremnts are between the same ranges and have similar distribution to the measuresd data, with slightly less irregularities.

